\newpage
\section{Mathematical Proofs}

In this section I will provide with some examples of different types of proofs.

\subsection{Proof by Direct Argument}

For any integer \(n\), if \(n\) is even, then \( n^2 \) is even.
We will prove this theorem by direct argument.

Assume \(n\) is an even integer. Then we can write \( n = 2k \) for some integer \(k\).

Now, we compute \( n^2 \):

\[		
	n^2 = {(2k)}^2 = 4k^2 = 2(2k^2)
\]
	
This shows that \( n^2 \) is even, as it can be expressed as \( 2m \) where \( m = 2k^2 \) is an integer.
Therefore, we conclude that if \(n\) is even, then \( n^2 \) is even.

\QED

\subsection{Proof by Contradiction}

If \(n\) is an integer such that \( n^2 \) is even, then \(n\) is even.

We will prove this theorem by contradiction. Assume that \(n\) is an integer such that \( n^2 \) is 
even, but \(n\) is odd. Then we can write \( n = 2k + 1 \) for some integer \(k\).

Now, we compute \( n^2 \):

\[
	n^2 = {(2k + 1)}^2 = 4k^2 + 4k + 1 = 2(2k^2 + 2k) + 1
\]
	
This shows that \( n^2 \) is odd, which contradicts our assumption that \( n^2 \) is even. Therefore, our 
assumption that \(n\) is odd must be false, and thus, \(n\) must be even.

\QED

Another example is the proof that there are infinite prime numbers.

Let us define the prime numbers \(p_1, \dots, p_n\) where we assume \(p_n\) is the biggest prime number. 
If we construct a number 

\[
	q = 1 + p_1 \cdot \cdots \cdot p_n 
\]

This number is bigger than \(p_n\) but it is not divisible by any of the other prime numbers will give us always a residual of 
1. Therefore the assumption that there is a biggest prime number is false.

\QED 

\subsection{Proof by Induction}

For all \( n \in \Naturals \), the sum of the first \(n\) positive integers is given by:

\[
	S(n) = 1 + 2 + 3 + \cdots + n = \frac{n(n+1)}{2}
\]

We will prove this theorem by induction on \(n\).

\textbf{Base Case:} For \( n = 1 \):

\[
	S(1) = 1 = \frac{1(1+1)}{2}
\]

The base case holds.

\textbf{Inductive Step:} Assume that the statement holds for some \( n = k \), i.e., assume that:
	
\[
	S(k) = 1 + 2 + 3 + \cdots + k = \frac{k(k+1)}{2}
\]
	
We need to show that the statement holds for \( n = k + 1 \):

\[
	S(k+1) = S(k) + (k + 1)
\]
	
By the inductive hypothesis, we have:

\begin{align*}
	S(k+1) &= \frac{k(k+1)}{2} + (k + 1) \\
	&= \frac{k(k+1)}{2} + \frac{2(k + 1)}{2}\\
    &= \frac{k(k+1) + 2(k + 1)}{2} \\
	&= \frac{(k + 1)(k + 2)}{2}	
\end{align*}

Thus, the statement holds for \( n = k + 1 \).
By the principle of mathematical induction, the statement holds for all \( n \in \Naturals \).

\QED

\subsection{Proof by Exhaustion}

The only integer solutions to the equation \( x^2 + y^2 = 1 \) are \( (0, 1), (1, 0), (0, -1), (-1, 0) \).

We will prove this theorem by exhaustion. We will check all possible integer values of \(x\) and 
\(y\) such that \( x^2 + y^2 = 1 \).

The possible integer values for \(x\) and \(y\) are \( -1, 0, 1 \). We will check each case:

If \( x = 0 \):

Then \( y^2 = 1 \) gives \( y = 1 \) or \( y = -1 \).

Solutions: \( (0, 1), (0, -1) \).

If \( x = 1 \):

Then \( y^2 = 0 \) gives \( y = 0 \).

Solution: \( (1, 0) \).

If \( x = -1 \):

Then \( y^2 = 0 \) gives \( y = 0 \).

Solution: \( (-1, 0) \).

Thus, the only integer solutions to the equation are:
	
\[
	(0, 1), (1, 0), (0, -1), (-1, 0)
\]

\QED

\subsection{Proof by Cases}

For any integer \(n\), \( n^2 \) is even if and only if \(n\) is even.

We will prove this theorem by cases.

\textbf{Case 1:} Assume \(n\) is even. Then we can write \( n = 2k \) for some integer \(k\).

Now, we compute \( n^2 \):

\[
	n^2 = {(2k)}^2 = 4k^2 = 2(2k^2)
\]

This shows that \( n^2 \) is even.

\textbf{Case 2:} Assume \(n\) is odd. Then we can write \( n = 2k + 1 \) for some integer \(k\).

Now, we compute \( n^2 \):

\[
	n^2 = {(2k + 1)}^2 = 4k^2 + 4k + 1 = 2(2k^2 + 2k) + 1
\]

This shows that \( n^2 \) is odd.

Since both cases have been considered, we conclude that \( n^2 \) is even if and only if \(n\) is even.

\QED

\subsection{Proof by Construction}

There exists an irrational number \(x\) such that \( x^2 \) is rational.

We will construct an irrational number \(x\) such that \( x^2 \) is rational.

Let \( x = \sqrt{2} \). We know that \( \sqrt{2} \) is irrational. Now, we compute \( x^2 \):
	
\[
	x^2 = (\sqrt{2})^2 = 2
\]
	
Since 2 is a rational number, we have constructed an irrational number \( x = \sqrt{2} \) such 
that \( x^2 = 2 \) is rational. Therefore, the theorem is proved.

\QED

\subsection{Proof by Counterexample}

The statement all prime numbers are odd is false.

To prove this theorem, we will provide a counterexample.

The number 2 is a prime number, as its only divisors are 1 and 2. However, 2 is 
even, which contradicts the statement that all prime numbers are odd.

Therefore, the statement All prime numbers are odd is false.
\QED

\subsection{Proof by Contrapositive}

If \(n\) is an integer such that \( n^2 \) is odd, then \(n\) is odd.

We will prove this theorem by contrapositive. The contrapositive of the statement is: If \(n\) is an 
integer such that \(n\) is even, then \( n^2 \) is even.

Assume \(n\) is even. Then we can write \( n = 2k \) for some integer \(k\).

Now, we compute \( n^2 \):
	
\[
	n^2 = {(2k)}^2 = 4k^2 = 2(2k^2)
\]

This shows that \( n^2 \) is even.

Since the contrapositive statement is true, the original statement If \( n^2 \) is odd, then \(n\) is 
odd is also true.

\QED

\subsection{Proof by Reduction to Absurdity}

The square root of 2 is irrational.

We will prove this theorem by reduction to absurdity. Assume that \( \sqrt{2} \) is rational. Then we can 
write:

\[
	\sqrt{2} = \frac{p}{q}
\]

Where \(p\) and \(q\) are integers with no common factors (i.e., the fraction is in the simplest 
form).

Squaring both sides gives:

\[
	2 = \frac{p^2}{q^2}
\]
	
Rearranging gives:

\[
	p^2 = 2q^2
\]
	
This implies that \( p^2 \) is even, and therefore, \(p\) must be even (since the square of an 
odd number is odd). Let \( p = 2k \) for some integer \(k\). Substituting this back into the 
equation gives:

\begin{align*}
{(2k)}^2 &= 2q^2\\	
4k^2 &= 2q^2\\
2k^2 &= q^2
\end{align*}

This implies that \( q^2 \) is even, and therefore, \(q\) must also be even.
Since both \(p\) and \(q\) are even, they have a common factor of 2, which contradicts our 
assumption that \(p\) and \(q\) have no common factors. Therefore, our assumption that 
\( \sqrt{2} \) is rational must be false, and thus, \( \sqrt{2} \) is irrational.

\QED

\subsection{Proof by Analogy}

The set of rational numbers is dense in the set of real numbers.

We will prove this theorem by analogy.

Consider the set of rational numbers \( \Rationals \) and the set of real numbers \( \Reals \). 
The density of \( \Rationals \) in \( \Reals \) means that between any two real numbers, there exists a 
rational number. For example, between the real numbers 1 and 2, we can find the rational 
number \( \frac{3}{2} = 1.5 \). Similarly, between any two real numbers \(a\) and \(b\) (where 
\( a < b \)), we can find a rational number \( r = \frac{a + b}{2} \).

This shows that the set of rational numbers is dense in the set of real numbers.
Therefore, the theorem is proved by analogy.

\QED

