\newpage
\section{Combinatorics}

\subsection{Three Principles of Combinatorics}

These principles are \emph{Addition}, \emph{Multiplication} and \emph{Inclusion-Exclusion}.

\subsection{Permutation}

Number of ways of ordering \(n\) distinct elements. In other words the number of tuples of n distinct elements.

\[
    n!
\]

\subsection{Permutation with repetition}

Number of ways of ordering \(n\) non-distinct elements. Here \(m_1, \dots, m_n\)
is the number a specific item is repeated in the original set. By dividing by the respective amount of the 
repeated non-distinct element, we eliminate the permutations, where the identical element swap their place.

\[
    \frac{n!}{m_1! \cdots m_n!}
\]

\subsection{Variation}

Ways of put \(n\) objects in \(k\) slots with repetition.

\[
    n^k
\]

\subsection{Variation without repetition}

Ways of put \(n\) objects in \(k\) slots without repetition. Number of tuples with \(k\) 
elements which can be selected from \(n\) elements, without repetition.

\[
    \frac{n!}{(n - k)!}
\]

\subsection{Combination without repetition}

Ways of choose \(k\) objects of \(n\) elements.

\[
    \binom{n}{k} = \frac{n!}{k!(n-k)!}
\]

Here \(n!\) is the number of permutations of the original set, the tuples.
\((n-k)!\) has the function of eliminating the permutations with more elements we are not interested in, and
\(k!\) has the function to eliminate the duplicates because for a combination the order does not matter, contrary to
the permutations. It is choose that way because for each choice of \(k\) objects there are \(k!\) possible orderings and we only 
want one.

\subsection{Combination with repetition}

This focuses not on the slots but in the separations between the slots.
More specific the number of ways to distribute \(k\) identical objects into \(n\) identical boxes. 

\[
    \binom{k + n - 1}{n - 1} = \binom{k + n - 1}{k} = \frac{(k + n - 1)!}{k!(n - 1)!}
\]

This comes from modeling our possible combinations. If we have a tuple \((a_1, a_2, \dots, a_n)\) where \(a_i\) gives us 
the number of occurrences of the object \(i\), then we can imagine using \(|\) a number of time for the occurrences of each object 
and \(\dot\) as the separator between each set of \(|\). For each possible combination we have \(k\) lines, \(k + n - 1\) 
positions, and \(n - 1\) points. Therefore, the question becomes How many ways are there of putting \(k\) lines into  \(k + n - 1\) positions.

\subsection{Multinomial Coefficients}

When using the binomial coefficient repeatedly one after the other we can abbreviate 
this consecutive uses as 

\[
    \binom{n}{k_1, \dots, k_n}
\]

This is like from \(n\) choose \(k_1\) then from \(n - k_1\) choose \(k_2\) and so on.

\subsubsection{Multinomial Theorem}

Given \(x_1, \dots, x_p \in \Reals\) and \(\Naturals\) then 

\[
    (x_1 + \cdots, x_n)^n = \sum_{k_1 + \cdots + k_{p = n}}  \binom{n}{k_1, \dots, k_p} x_{1}^{k_1} \cdots x_{p}^{k_p}
\]

\subsection{Disarray}

Number of permutation in which no object ends in the same initial spot. Also, know as the 
\emph{sub-factorial}

\[
    !n = n! \sum_{k = 0}^{n} \left(\frac{{(-1)}^k}{k!}\right)
\]

\subsection{Bell Numbers}

Number of ways of grouping \(n\) objects in an arbitrary number of slots.

\[
    B(N) = \sum_{k = 0}^{N-1}\binom{N}{K}B(K)
\]

with \(B(1) = 1\) and \(B(2) = 2\)

\subsection{Ramanujan Numbers}

The same as the Bell Numbers but, all objects are equal. As an example the number of ways
to decompose a number.

\[
    R(N) \approx \frac{1}{4N\sqrt{3}} e^{2\pi \sqrt{\frac{N}{6}}}
\]

\subsection{Stirling Numbers I}

Ways of permuting \(N\) items with \(K\) exchanges.

\[
    \genfrac{[}{]}{0pt}{}{N + 1}{K} = N \genfrac{[}{]}{0pt}{}{N}{K} + \genfrac{[}{]}{0pt}{}{N}{K - 1}
\]

\subsection{Stirling Numbers II}

Ways of grouping \(n\) items in \(k\) slots.

\[
    \genfrac{\{}{\}}{0pt}{}{N}{K} = \frac{1}{K!} \sum_{j=0}^{K} (-1)^{K - j} \binom{K}{j} j^N
\]

\subsection{Lah Numbers}

Ways of building \(K\) lists with a set \(N\) elements.

\[
    \genfrac{\lfloor}{\rfloor}{0pt}{}{N}{K} = \binom{N - 1}{K - 1} \frac{N!}{K!} = \sum_{j=0}^{K} 
    \genfrac{[}{]}{0pt}{}{N}{K} \genfrac{\{}{\}}{0pt}{}{N}{K}
\]

\subsection{Euler's Numbers}

Number of permutations of a set of \(N\) elements of different size in which
\(K\) elements are bigger than their previous element.

\[
    \genfrac{\langle}{\rangle}{0pt}{}{N}{K} = \sum_{j = 0}^{K} (-1)^j \binom{N + 1}{j} (K + 1 - j)^N
\]

\subsection{Catalan Numbers}

These have different applications like the number of ways of the triangulations of
a polygon or the number of trees with \(N\) leafs.

\[
    \mathfrak{C}(N) = \binom{2N}{N} - \binom{2N}{N + 1}
\]

or

\[
    \mathfrak{C}(N) = \sum_{k=1}^{n}(k - 1) \mathfrak{C}(N - k)
\]

\subsection{Pascals Triangle}

This is a triangle build by the formula \(\binom{n + 1}{k + 1} = \binom{n}{k}\binom{n}{k + 1}\)

\begin{center}
    \begin{tabular}{>{$n=}l<{$\hspace{12pt}}*{13}{c}}
    0 &&&&&&&1&&&&&&\\
    1 &&&&&&1&&1&&&&&\\
    2 &&&&&1&&2&&1&&&&\\
    3 &&&&1&&3&&3&&1&&&\\
    4 &&&1&&4&&6&&4&&1&&\\
    5 &&1&&5&&10&&10&&5&&1&\\
    6 &1&&6&&15&&20&&15&&6&&1
    \end{tabular}
\end{center}
\vspace{1cm}

Here is the triangle but with the Binomial Coefficients
\smallskip

\begin{center}
    \begin{tabular}{>{$n=}l<{$\hspace{12pt}}*{13}{c}}
    0 &&&&&&&\(\binom{0}{0}\)&&&&&&\\
    1 &&&&&&\(\binom{1}{0}\)&&\(\binom{1}{1}\)&&&&&\\
    2 &&&&&\(\binom{2}{0}\)&&\(\binom{2}{1}\)&&\(\binom{2}{2}\)&&&&\\
    3 &&&&\(\binom{3}{0}\)&&\(\binom{3}{1}\)&&\(\binom{3}{2}\)&&\(\binom{3}{3}\)&&&\\
    \end{tabular}
\end{center}

It can be used with the Binomial Theorem to get the Coefficients of a polynomial of degree \(n\).

\textbf{Example:}

\[
    {(a + b)}^2 = 1 a^2b^0 + 2a^1b^1 + 1a^0b^2
\]

or

\[
    {(a - b)}^2 = 1 a^2b^0 - 2a^1b^1 + 1a^0b^2
\]
