\newpage
\section{Continuity}

Continuity describes functions whose values change smoothly, without abrupt jumps or breaks. 
This concept is fundamental in calculus, analysis, and fixed point theory.

\subsection{Definition of Continuity at a Point}

A function \(f\) is \emph{continuous} at a point \(x_0\) if:

\[
    \forall \varepsilon > 0, \exists \delta > 0 \text{ such that } |x - x_0| < \delta \Rightarrow |f(x) - f(x_0)| < \varepsilon
\]

Equivalently,

\[
    \lim_{x \to x_0} f(x) = f(x_0)
\]

To solve continuity problems the general approach is to:

\begin{enumerate}

    \item Write down the definition.

    \item Substitute \(f\) for the concrete function into the definition.

    \item Modify the expression with the concrete \(f\) \(|f(x) - f(y)|\) in way that you get 
          \(| x - y|\) inside the expression.
    
    \item Choose a \(\delta\) if not given. On the one side it is often \(\delta = |x - y |\) and then 
          substitute into the expression. On the other side if \(\delta\) is given make an approximation 
          which is greater than the current expression and then simplify. In both cases after simplifying 
          you should get a \(\varepsilon\) is the function is continuous. 

\end{enumerate}

\textbf{Example:} 

\(\varepsilon-\delta\) Exercise for \(f(x) = \frac{1}{x}\) at \(x_0 > 0\)

We want to find \(\delta\) such that:

\[
    |x - x_0| < \delta \Rightarrow \left|\frac{1}{x} - \frac{1}{x_0}\right| < \varepsilon
\]

Start by rewriting:
\[
    \left|\frac{1}{x} - \frac{1}{x_0}\right| = \left|\frac{x_0 - x}{xx_0}\right| = \frac{|x - x_0|}{|x||x_0|}
\]

Assume:

\[
    x \in \left[\frac{x_0}{2}, \frac{3x_0}{2}\right] \Rightarrow |x| \ge \frac{x_0}{2}
\]

Then:

\[
    \left|\frac{1}{x} - \frac{1}{x_0}\right| \le \frac{|x - x_0|}{\frac{x_0}{2} \cdot x_0} = \frac{2}{x_0^2} |x - x_0|
\]

Choose:

\[
    \delta = \min\left\{\frac{x_0}{2}, \frac{x_0^2 \varepsilon}{2} \right\}
\]

\subsection{Rules of Continuity}

Let \(f\) and \(g\) be continuous at \(x_0\):

\begin{itemize}
    
    \item \(f + g\), \(f - g\), \(f \cdot g\) are continuous at \(x_0\)
    
    \item \(\frac{f}{g}\) is continuous at \(x_0\) if \(g(x_0) \ne 0\)
    
    \item Compositions: if \(g\) is continuous at \(x_0\), and \(f\) is continuous at \(g(x_0)\), 
    then \(f \circ g\) is continuous at \(x_0\)

\end{itemize}

\subsection{Lipschitz Continuity}

A function \(f: D \subset \Reals \to \Reals\) is \emph{Lipschitz continuous} if:

\[
    \exists L > 0 \text{ such that } |f(x) - f(y)| \le L |x - y| \quad \forall x, y \in D
\]

\begin{itemize}

    \item Every Lipschitz continuous function is uniformly continuous.

    \item The smallest such \(L\) is called the \emph{Lipschitz constant}.

\end{itemize}

\subsection{Banach Fixed Point Theorem}

Let \((X, d)\) be a complete metric space, and \(f: X \to X\) a \emph{contraction mapping}, i.e.,

\[
    \exists L < 1 \text{ such that } d(f(x), f(y)) \le L \cdot d(x, y)
\]

Then:

\begin{itemize}

    \item \(f\) has a unique fixed point \(x^* \in X\), such that \(f(x^*) = x^*\)

    \item Iterating \(x_{n+1} = f(x_n)\) converges to \(x^*\)

\end{itemize}

\emph{Interpretation and Meaning}

The Banach Fixed Point Theorem ensures:

\begin{itemize}

    \item Existence and uniqueness of solutions (fixed points)

    \item Convergence of approximation by iteration

    \item Powerful tool in numerical methods and differential equations

\end{itemize}

\subsubsection{A Priori and A Posteriori Approximations}

\emph{A priori estimate:}

\[
    |x_n - x^*| \le \frac{L^n}{1 - L} |x_1 - x_0| < \epsilon
\]

\emph{A posteriori estimate:}

\[
    |x_n - x^*| \le \frac{L}{1 - L} |x_n - x_{n-1}| \epsilon
\]

\subsubsection{Steps for a Fixed Point Exercise}

Given \(f: [a, b] \to [a, b]\), to prove existence and convergence:

\begin{enumerate}

    \item Show monotonicity: \(f\) is increasing or decreasing.

    \item Check interval preservation: \(f([a, b]) \subseteq [a, b]\)

    \item Check Lipschitz continuity: Find \(L < 1\)

    \item Fixed point iteration: Choose \(x_0\), compute \(x_{n+1} = f(x_n)\)

    \item Apply a priori or a posteriori bound

\end{enumerate}

\textbf{Example:} 

\(f(x) = \frac{1}{x} + 3\) on \([2, 5]\)

\textbf{Step 1: Monotonicity}

In this case we will assume that it is monotonous. 

\textbf{Step 2: Domain check} 
 
 \(x \in [2, 5] \Rightarrow \frac{1}{x} \in [0.2, 0.5] \Rightarrow f(x) \in [3.2, 3.5] \subset [2, 5]\)

 \textbf{Step 3: Lipschitz constant}

\[
    f'(x) = -\frac{1}{x^2} \Rightarrow |f'(x)| \le \frac{1}{2^2} = \frac{1}{4} < 1
    \Rightarrow \text{Lipschitz with } L = \frac{1}{4} < 1
\]


\textbf{Step 4: Fixed point iteration}

\[
    x_0 = 3, \quad x_1 = f(3) = \frac{1}{3} + 3 = \frac{10}{3}, \quad x_2 = f(x_1), \dots
\]

\textbf{Step 5: Approximations}

\emph{A priori}

\[
    |x_n - x^*| \le \frac{L^n}{1 - L} |x_1 - x_0|
\]

\[
    |x_n - x^*| \le \frac{\frac{1}{4}^n}{1 - \frac{1}{4}} |\frac{10}{3}  - 3|
\]

\begin{align*}
    |x_n - x^*| &\le \frac{\left(\frac{1}{4}\right)^n}{1 - \frac{1}{4}} \cdot \left|\frac{10}{3} - 3\right| \\
    &= \frac{\left(\frac{1}{4}\right)^n}{\frac{3}{4}} \cdot \frac{1}{3}\\
    &= \frac{4}{3} \cdot \left(\frac{1}{4}\right)^n \cdot \frac{1}{3}\\
    &= \frac{4}{9} \cdot \left(\frac{1}{4}\right)^n
\end{align*}

For \(\varepsilon = 0,1\)

\begin{align*}
    \frac{4}{9} \cdot \left(\frac{1}{4}\right)^n &< \varepsilon \\
    \left(\frac{1}{4}\right)^n &< \frac{9}{4} \varepsilon \\
    \ln\left(\left(\frac{1}{4}\right)^n\right) &< \ln\left(\frac{9}{4} \varepsilon\right) \\
    n \cdot \ln\left(\frac{1}{4}\right) &< \ln\left(\frac{9}{4} \varepsilon\right) \\
    n &> \frac{\ln\left(\frac{9}{4} \varepsilon\right)}{\ln\left(\frac{1}{4}\right)} \\
    n &> \frac{\ln\left(\frac{9}{4} \varepsilon\right)}{-\ln 4}
\end{align*}


\emph{A posteriori}

\[
    |x_n - x^*| \le \frac{L}{1 - L} |x_n - x_{n-1}| = \frac{\frac{1}{4}}{1 - \frac{1}{4}} |x_n - x_{n-1}| = \frac{1}{3} |x_n - x_{n-1}|
\]

