\newpage
\section{Topology}

In this section, we introduce essential vocabulary used in topology. Each term is accompanied by a 
brief explanation and its formal mathematical definition.

\subsection{Introduction to topological nomenclature}

\subsubsection{Open Ball \( B_\varepsilon(x) \)} 

For a metric space \( (X, d) \), the \emph{open ball} centered at 
\( x \in X \) with radius \( \varepsilon > 0 \) is defined as: 
	      
\[
	B_\varepsilon(x) := \{ y \in X \mid d(x, y) < \varepsilon \} = (x - \varepsilon, x + \varepsilon)
\]

It represents the set of all points within distance \( \varepsilon \) from \(x\), excluding the boundary.

\subsubsection{Open Set}
	     
A subset \( U \subseteq X \) of a topological space is called 
\emph{open} if for every point \( x \in U \), there exists an \( \varepsilon > 0 \) 
such that the open ball \( B_\varepsilon(x) \subseteq U \). 
Intuitively, an open set contains none of its boundary points and every point has some 
wiggle room around it.

\textbf{Example:}

We will prove that  \((0,1)\) is an open set. 

This means \(x_0 > 0\) and \(1 - x_0 > 0\). First we define 

\[
	\varepsilon := \frac{1}{2} \min {x_0, 1 - x_0}
\]

Then from \(|y - x_0| \le \varepsilon\) we get 

\begin{align*}
	&y \ge x_0 |y - x_0 | \ge x_0 - \varepsilon \ge \frac{1}{2}x_0 > 0, \\ 
	&y \le x_0 |y - x_0 | \le x_0 + \varepsilon \le x_0 + \frac{1}{2}(1 - x_0) = \frac{1}{2} + \frac{1}{2}x_0 < 1
\end{align*}

\QED 

\subsubsection{The Interior of a Set} 

The \emph{interior} denoted as \(A^{\circ}\) of a set \(A\) is the set of all inner points of \(A\). 

\subsubsection{Closed Set} 
	      
A subset \( A \subseteq X \) is called \emph{closed} if its 
complement \( X \setminus A \) is open. Equivalently, \(A\) contains all its limit points. 
That is, \(A\) is closed if it includes its boundary.

\textbf{Example:} \([0,1]\)

\subsubsection{Boundary Point}
	      
A point \( x \in X \) is a \emph{boundary point} of a set \( A \subseteq X \) 
if every open ball around \(x\) contains both points in \(A\) and in \( X \setminus A \). 
The set of all boundary points is called the \emph{boundary} of \(A\), denoted \( \partial A \).

\subsubsection{Accumulation Point / Limit Point}

A point \( x \in X \) is an \emph{accumulation point} of a set \( A \subseteq X \)
if every open ball \( B_\varepsilon(x) \) contains a point of \( A \setminus \{x\} \). 
In other words, points of \(A\) cluster arbitrarily close to 
\(x\), even if \( x \notin A \).

\subsubsection{Isolated Point} 
	      
A point \( x \in A \) is an \emph{isolated point} if there exists \( \varepsilon > 0 \)
such that \( B_\varepsilon(x) \cap A = \{x\} \). 
That is, \(x\) stands alone in \(A\) without other points of \(A\) nearby.

\textbf{Example:}

\(-1\) is an isolated point of \({-1} \cup (0,1)\)

\subsubsection{Boundary}

The set \(A \cup \{\text{ all boundary points }\}\) is called the \emph{boundary} of \(A\). It is denoted 
as \(\overline{A}\).

\subsubsection{Compact Set} 

A set \( K \subseteq X \) is \emph{compact} if every open cover of \( K \) has a finite sub-cover. 
In \(\Reals^n\), this is equivalent to \( K \) being closed and bounded (by the Heine–Borel theorem).

\subsubsection{Dense Set} 
	      
A subset \( D \subseteq X \) is \emph{dense} in \(X\) if every point 
\( x \in X \) is either in \( D \) or is a limit point of \( D \). 
Equivalently, the closure of \( D \) is \(X\), i.e., \( \overline{D} = X \).


